Ab Initio Effective One-Electron Potential Operators. I. Applications for Charge-Transfer Energy in Effective Fragment Potentials

### 2019-12-21 papers/2_oep_technique.CT/paper-r4.tex
Revision 4 of Paper 2 added

\documentclass[aip,jcp,amsmath,amssymb,reprint,floatfix]{revtex4-1}
\input{header.tex}
\begin{document}
\preprint{AIP/123-OEP}

\title{Ab Initio Effective One-Electron Potential Operators:
Applications for Charge-Transfer Energy in Effective Fragment Potentials}

@@

\begin{abstract}
The concept of the effective one\hyp{}electron potentials (OEP) has been useful for many decades
in efficient description of electronic structure of chemical systems, especially extended
molecular aggregates such as interacting molecules in condensed phases.
Here, a general method for effective OEP\hyp{}based elimination of electron repulsion integrals (ERI),
that is tuned
towards the fragment\hyp{}based calculation methodologies
such as the second generation of the effective fragment potentials (EFP2) method,
is presented.
Two general types of the OEP operator matrix elements
are distinguished and treated either via the distributed multipole expansion
or the extended density fitting schemes developed in this work. 
The OEP technique is then applied
to address the problem of using incomplete EFP2 settings in many
applications in interaction energy and 
molecular dynamics simulations 
due to prohibitively high computational cost of evaluating the
charge transfer (CT) effects as compared to other effects.
The alternative OEP\hyp{}based CT energy model is proposed
in the context of the
intermolecular perturbation theory with Hartree\hyp{}Fock uninteracting gas\hyp{}phase 
reference wavefunctions, compatible with the EFP2 formulation.
It is found that the computational cost can be reduced
up to 20 times as compared to the CT energy method within the EFP2 scheme
without compromising the accuracy for a wide range of weakly interacting 
neutral molecular complexes. 
Therefore, it is believed that
the proposed model can be used within 
the EFP2 framework, making the CT energy term no longer the bottleneck
in EFP2\hyp{}based simulations of complex systems.
\end{abstract}

@@

The apparent difficulty in theoretically characterising the CT energy
in terms of the interacting molecular fragments
is indeed a challenge in developing modern force fields 
or \emph{ab initio} fragmentation methods\cite{Gordon.Fedorov.Pruitt.Slipchenko.ChemRev.2012}
for modeling structure
and dynamics in condensed phases.\cite{Demerdash.Mao.Liu.Head-Gordon.Head-Gordon.JCP.2017}
This needs to be contrasted with the Coulombic electrostatics,
non\hyp{}Coulombic repulsion (i.e., due to Pauli exclusion principle), 
dispersion and induction for instance, 
which to a certain extent
can be well described by, respectively, 
the distributed multipole moments of charge densities,\cite{Sokalski.Poirier.CPL.1983,
Etchebest.Lavery.Pullman.TheorChimActa.1982,Stone.JCTC.2005}
the van der Waals repulsive and attractive potentials, 
as well as the polarizability models, i.e., all being relatively simple and computationally
inexpensive to evaluate mathematical models. 
Due to this reason, the CT effects 
%while usually of significant and non\hyp{}negligible magnitude as compared 
%to the total interaction energy, 
are not explicitly included in any molecular mechanics force field 
developed up to date.\cite{Demerdash.Yap.Head-Gordon.AnnuRevPhysChem.2014}
%such as OPLS, Amber, CHARMM, Gromos or even polarizale AMOEBA force field. 
The only way to
incorporate the CT effects in the condensed\hyp{}phase simulations
in \emph{ab initio} manner,
apart from performing full QM electronic structure simulations
or non\hyp{}force\hyp{}field\hyp{}based 
fragmentation techniques\cite{Leverentz.Maerzke.Keasler.Siepmann.Truhlar.PCCP.2012,
Dahlke.Truhlar.JCTC.2008},
is through the second generation of the effective fragment potential 
(EFP2) method,\cite{Gordon.Smith.Xu.Slipchenko.AnnuRevPhysChem.2013,
   Nguyen.Pachter.Day.JCP.2014,
   Day.Jensen.Gordon.Webb.Stevens.Krauss.Garmer.Basch.Cohen.JCP.1996,
   Sattasathuchana.Xu.Gordon.JPCA.2019,
   Blasiak.Londergan.Webb.Cho.ACR.2017,
   Kuroki.Mori.ChemLett.2016,
   Ghosh.Cho.Choi.JPCB.2014}
developed by Jensen, Gordon and co\hyp{}workers.\cite{Jensen.JCP.2001,Li.Gordon.Jensen.JCP.2006,Xu.Gordon.JCP.2013}
EFP2 is an \emph{ab initio} force field
derived on the grounds of the first\hyp{}principles %perturbation theory
at the HF level\cite{Jensen.Gordon.MolPhys.1996,Jensen.Gordon.JCP.1998,Li.Netzloff.Gordon.JCP.2006} 
and including the intermolecular dispersion effects
by the response theory.\cite{Adamovic.Gordon.MolPhys.2005,Xu.Zahariev.Gordon.JCTC.2014}
%based on the imaginary frequency\hyp{}dependent dipole\hyp{}dipole polarizabilities.
That is to say, the total intermolecular 
interaction potential, which can also be perceived as an approximation
to SAPT at HF level with added intermolecular dispersion correction, is approximated as 
%
\begin{equation}\label{e:u-efp2}
 U^{\rm EFP2} \approx U^{\rm Coul} + U^{\rm Ex-Rep} + U^{\rm Ind} + U^{\rm Disp} + U^{\rm CT} \;,
\end{equation}
%
where $U^{\rm Coul}$ is the Coulombic interaction energy of the unperturbed charge\hyp{}density
distributions of the monomers, treated by the distributed multipole approximation
with damping to account for the charge\hyp{}penetration effects,\cite{Slipchenko.Gordon.JCC.2007}
$U^{\rm Ex-Rep}$ is the exchange\hyp{}repulsion energy originating from the Pauli exclusion
principle,\cite{Jensen.Gordon.MolPhys.1996,Jensen.Gordon.JCP.1998} 
that could be roughly understood as the Heitler\hyp{}London interaction
energy\cite{Chalasinski.Gutowski.MolPhys.1985} with $U^{\rm Coul}$ subtracted, $U^{\rm Ind}$ and $U^{\rm Disp}$
are the induction and dispersion energies obtained from the distributed polarizability
approximation,\cite{Li.Netzloff.Gordon.JCP.2006,Adamovic.Gordon.MolPhys.2005,Xu.Zahariev.Gordon.JCTC.2014} 
and finally $U^{\rm CT}$ is the CT energy,\cite{Li.Gordon.Jensen.JCP.2006,Xu.Gordon.JCP.2013} 
the focus of this work.

@@

To establish the notation within the present work, 
the formalism of one\hyp{}electron potential operators is
given first.
The one\hyp{}electron Coulomb static effective potential $v^{\rm eff}({\bf r})$
produced by a certain effective one\hyp{}electron charge density distribution $\rho^{\rm eff}({\bf r})$
is 
%
\begin{equation} \label{e:v-eff}
	v^{\rm eff}({\bf r}) = \int \frac{ \rho^{\rm eff}({\bf r}') }{ \vert {\bf r}' - {\bf r} \vert} d{\bf r}' \;,
\end{equation}
%
where ${\bf r}$ is a spatial coordinate. 
For convenience, the total effective density can be split into nuclear and electronic contributions,
%
\begin{equation} \label{e:rho-eff}
 \rho^{\rm eff}({\bf r}) = \lambda \rho^{\rm eff}_{\rm nuc}({\bf r}) + \rho^{\rm eff}_{\rm el}({\bf r}) \;,
\end{equation}
%
where $\lambda$ is a certain parameter and is assumed to be either 1 or 0 in this work.
Formally, the electronic part can be expanded in terms of an effective
bond order matrix, or one\hyp{}particle density (OED), represented in a certain basis of orbitals, $\BM{\phi}({\bf r})$,
%
\begin{equation} \label{e:d-spectral}
	\rho^{\rm eff}_{\rm el}({\bf r}) = \sum_{\alpha\beta} P_{\alpha\beta}^{\rm eff} 
	\phi_\alpha({\bf r}) \phi_\beta^{*}({\bf r})  \;.
\end{equation}
%
Based on that, the operator form of the effective potential 
can be written as
%
\begin{equation} \label{e:oep-operator}
	\hat{v}^{\rm eff} = 
        \lambda \hat{v}_{\rm nuc} +
        %\sum_{\alpha\beta} 
	%\Ket{\alpha} 
        \int d{\bf r} \Ket{{\bf r}} 
        v^{\rm eff}_{\rm el}({\bf r})
	%\left[
	%\int
	%\hat{J}^{\rm}_{\alpha\beta}({\bf r}') 
	%\rho^{\rm eff}({\bf r}')
	%d{\bf r}' 
	%\right]
	%\Bra{\beta}  
        \Bra{{\bf r}}
\end{equation}
%
with the nuclear
%and Coulomb operators 
operator
defined by
%
%\begin{subequations}
%\begin{align}
\begin{equation} \label{e:oep-operator-nuc}
        \hat{v}_{\rm nuc} \equiv \sum_x^{\rm At}  
                     \int d{\bf r} \Ket{{\bf r}} 
                     \frac{Z_x}{\vert {\bf r} - {\bf r}_x \vert}
                     \Bra{{\bf r}} \;.
	%\hat{J}^{\rm}_{\alpha\beta}({\bf r}) f({\bf r}) &\equiv
	%f({\bf r}) 
	%\int
	%\frac{ \phi_\alpha^{*}({\bf r}') \phi_\beta({\bf r}') }{ \vert {\bf r}' - {\bf r} \vert} d{\bf r}'
\end{equation}
%\end{align}
%\end{subequations}
%
%for any one\hyp{}electron function $f({\bf r})$. 
and $\BraKet{\bf r}{\alpha} \equiv \varphi_\alpha({\bf r})$.
The matrix element of the effective potential operator
is therefore given by
%
\begin{equation} \label{e:oep-matrix-element}
	\tBraKet{\alpha}{ \hat{v}^{\rm eff} }{\beta}
	= \lambda \sum_x^{\rm At} W_{\alpha\beta}^{(x)} +
        \sum_{\gamma\delta} \BraKet{\alpha\beta}{\gamma\delta} P^{\rm eff}_{\gamma\delta}  \;,
\end{equation}
%
where 
%
\begin{equation} \label{e:oep-matrix-element-nuc}
 W_{\alpha\beta}^{(x)} = 
 Z_x \int \frac{\phi^*_\alpha({\bf r}) \phi_\beta({\bf r})}{\vert {\bf r} - {\bf r}_x \vert} d{\bf r} \;,
\end{equation}
%
$Z_x$ is the atomic number of the $x$th atom,
and the symbol `At' denotes all atoms that contribute to the effective potential.
In the above equation and throughout the work, 
%
\begin{equation} \label{e:1el-operator}
\tBraKet{\alpha}{\mathscr{O}(1)}{\beta} \equiv \int \phi^*_\alpha({\bf r}) \mathscr{O}(1) \phi_\beta({\bf r}) d{\bf r} 
\end{equation}
%
for any one\hyp{}electron operator $\mathscr{O}(1)$ 
and the ERI
is defined according to
%
\begin{equation} \label{e:eri}
	\BraKet{\alpha\beta}{\gamma\delta} \equiv
	\iint 
	\frac{ \phi_\alpha^{*}({\bf r}_1) \phi_\beta({\bf r}_1) 
	       \phi_\gamma^{*}({\bf r}_2) \phi_\delta({\bf r}_2) }{ \vert {\bf r}_1 - {\bf r}_2 \vert}
	d{\bf r}_1 d{\bf r}_2  \;.
\end{equation}
%

@@

Three classes of ERI's can be recognized 
based on the basis function partitioning scheme
within the system composed of two molecules (shall be $A$ and $B$
throughout the course of this work),
in which it is possible to define an OEP. 
They are as follows:
%
\begin{enumerate}
\item the Coulomb\hyp{}like ERI's of the type 
$\BraKet{AA}{BB}\rightarrow \BraKet{\phi_{i\in A}\phi_{k\in A}}{\phi_{j\in B}\phi_{l\in B}}$,
\item the overlap\hyp{}like ERI's of the type 
$\BraKet{AA}{AB}\rightarrow \BraKet{\phi_{i\in A}\phi_{k\in A}}{\phi_{j\in A}\phi_{l\in B}}$, and
\item the exchange\hyp{}like ERI's of the type 
$\BraKet{AB}{AB}\rightarrow \BraKet{\phi_{i\in A}\phi_{j\in B}}{\phi_{k\in A}\phi_{l\in B}}$.
\end{enumerate}
%
Only the first two classes of ERI's can be incorporated into OEP's,
while exchange\hyp{}like ERI's cannot.
The two groups, which are listed in Table~\ref{t:oep-matrix-element-types},
are usually approximated via
expanding the OEP operator in distributed multipole (DMTP) expansion series, and integrating over
one remaining electron coordinate,\cite{Li.Gordon.Jensen.JCP.2006} i.e.,
%
\begin{multline} \label{e:coulomb-energy:dmtp-oei}
 \tBraKet{j}{\hat{v}_{\rm eff}^A}{l} \cong \\
 \left< j \left| 
                   \sum_{w\in A} \left\{
                                 T^{(0)}_{wu} \hat{q}_w 
                          - {\bf T}^{(1)}_{wu} \cdot \hat{\BM\upmu}_w 
            + \frac{1}{3}   {\bf T}^{(2)}_{wu} :     \hat{\BM\Theta}_w 
                    - \ldots  \right\}
              \right| l \right> \;.
\end{multline}
%
In the above equation, $\hat{q}_{\rm eff}^{(w)}$,
$\hat{\BM\upmu}^{(w)}_{\rm eff}$ and $\hat{\BM\Theta}^{(w)}_{\rm eff}$
are the quantum operators of effective distributed monopole (charge),
dipole moment and quadrupole moment, respectively, centered on the $w$th site, 
whereas ${{\bf T}^{(r)}_{wu}}$ are the so called interaction tensors of rank $r$ 
that can be found elsewhere.\cite{Blasiak.Cho.JCP.2014}
Note that $\Ket{j}$ and $\Ket{l}$ can belong to either molecule.
In this way, ERI's are no longer needed and the computational cost reduces appreciably.
In this work, this method is however considered already too expensive
for application in the CT/HF0 energy because evaluation of
Eq.~\eqref{e:coulomb-energy:dmtp-oei} requires calculation
of electrostatic potential and electrostatic potential gradient(s) OEI's. These kind of
integrals are typically the most expensive as compared to other standard OEI's
such as overlap or kinetic energy integrals.
Therefore, Coulomb and overlap\hyp{}like OEP matrix elements
%The first type is a Coulomb\hyp{}like contribution, for which 
will be treated via more approximate and less expensive approaches
%the multipole expansion or density fitting
which are discussed next.
% can be used.
%The second type is an overlap\hyp{}like contribution, which can be approximated via the density fitting. 
%Note that the exchange\hyp{}like integrals cannot be treated with the OEP technique.

%
{
\renewcommand{\arraystretch}{1.4}
\begin{table}[b]
\caption[Types of matrix elements with OEP operators]
{{\bf Types of matrix elements with OEP operators\footnotemark[1]}
}
\label{t:oep-matrix-element-types}
\begin{ruledtabular}
\begin{tabular}{lcccc}
Matrix element      &&            `Overlap-like'                &&            `Coulomb-like'               \\ 
                    && $\tBraKet{i}{\hat{v}^{A}_{\rm eff}}{j} $ && $\tBraKet{j}{\hat{v}^{A}_{\rm eff}}{l}$ \\ 
	\cline{1-5}
Partitioning scheme &&            $i\in A, j\in B$              &&               $j,l\in B$                \\
ERI class           &&            $\BraKet{AA}{AB}$             &&               $\BraKet{AA}{BB}$         \\
DF\footnotemark[2]/RI\footnotemark[3] Form    
&& $\sum_{\xi\in A}^{\rm Aux} v^A_{i\xi} S^{AB}_{\xi j} $  
&& $\sum_{\xi\zeta\in A}^{\rm Aux} S^{BA}_{j\xi} v^A_{\xi\zeta} S^{AB}_{\zeta l} $ \\
DMTP\footnotemark[4] Form                     
&& --  &  &  $\rho_{jl}^B \odot \rho_{\rm eff}^A$ \\
\end{tabular}
\end{ruledtabular}
%
\footnotetext[1]{Test footnotemark.}
\footnotetext[2]{Density Fitting}
\footnotetext[3]{Resolution of Identity}
\footnotetext[4]{Distributed Multipole Expansion}
%
\end{table}
}
%

@@

This scheme, which will be referred to as the EDF scheme, 
is applicable in case of matrix elements of 
$
 \tBraKet{i^A}{\hat{v}_{\rm eff}^A}{j^B}
$ type. 
 These matrix elements require ERI's of type
$\BraKet{AA}{AB}$ only.
To get the \emph{ab initio} representation of such an overlap\hyp{}like matrix element,
one can use a procedure similar to
the typical density fitting or resolution of identity, both of which are nowadays widely used 
to compute electron\hyp{}repulsion integrals (ERI's) more efficiently,
and reduce computational cost of post\hyp{}Hartree\hyp{}Fock methods. Density fitting was also
applied to design \emph{ab initio} force fields.\cite{Piquemal.Cisneros.Reinhardt.Gresh.Darden.JCP.2006,Cisneros.Andres.Piquemal.Darden.JCP.2005}

@@

the following objective function

@@

by performing integrations in Eq.~\eqref{e:triple-integral} is very costly 

@@

$H_{i\varepsilon}$ are given by Eq.~\eqref{e:v-compl}.
Thus, the matrix elements of the OEP operator can be found

@@

$V^A_{i\xi}$ can be calculated just once and stored in a file as effective fragment parameters.

Since, in contrast to the EDF-1 scheme, now two density fitting steps are required to
obtain OEP matrix, it will be referred to as the EDF-2 scheme.
In the limitting case, when the auxiliary basis is the same as the intermediate basis,
EDF-2 is equivalent to EDF-1.
As the theory required to define OEP's and their computational
evaluation in the fragment\hyp{}based formulation has been finally given in this Section,
the application of the OEP technique of ERI elimination is
further discussed for the evaluation of the CT/HF0 energy in the next sections.

@@

 \Bra{i} \left[ \hat{v}^B_{nj} \Ket{j} \right]  \;,

>>>>

\documentclass[aip,jcp,amsmath,amssymb,reprint,floatfix]{revtex4-1}
\input{header.tex}
\begin{document}
%\preprint{AIP/123-OEP}

\title{Ab Initio Effective One-Electron Potential Operators. I.
Applications for Charge-Transfer Energy in Effective Fragment Potentials}

@@

\begin{abstract}
The concept of the effective one\hyp{}electron potentials (OEP) has been useful for many decades
in efficient description of electronic structure of chemical systems, especially extended
molecular aggregates such as interacting molecules in condensed phases.
Here, a general method for effective OEP\hyp{}based elimination of electron repulsion integrals (ERI),
that is tuned
towards the fragment\hyp{}based calculation methodologies
such as the second generation of the effective fragment potentials (EFP2) method,
is presented.
Two general types of the OEP operator matrix elements
are distinguished and treated either via the distributed multipole expansion
or the extended density fitting schemes developed in this work. 
The OEP technique is then applied
to address the problem of using incomplete EFP2 settings in many
applications in interaction energy and 
molecular dynamics simulations 
due to relatively high computational cost of evaluating the
charge transfer (CT) effects as compared to other effects.
The alternative OEP\hyp{}based CT energy model is proposed
in the context of the
intermolecular perturbation theory with Hartree\hyp{}Fock uninteracting gas\hyp{}phase 
reference wavefunctions, compatible with the EFP2 formulation.
It is found that the computational cost can be reduced
up to 20 times as compared to the CT energy method within the EFP2 scheme
without compromising the accuracy for a wide range of weakly interacting 
neutral molecular complexes. 
Therefore, it is believed that
the proposed model can be used within 
the EFP2 framework, making the CT energy term no longer the bottleneck
in EFP2\hyp{}based simulations of complex systems.
\end{abstract}

@@

The apparent difficulty in theoretically characterising the CT energy
in terms of the interacting molecular fragments
is indeed a challenge in developing modern force fields 
or \emph{ab initio} fragmentation methods\cite{Gordon.Fedorov.Pruitt.Slipchenko.ChemRev.2012}
for modeling structure
and dynamics in condensed phases.\cite{Demerdash.Mao.Liu.Head-Gordon.Head-Gordon.JCP.2017}
This needs to be contrasted with the Coulombic electrostatics,
non\hyp{}Coulombic repulsion (i.e., due to Pauli exclusion principle), 
dispersion and induction for instance, 
which to a certain extent
can be well described by, respectively, 
the distributed multipole moments of charge densities,\cite{Sokalski.Poirier.CPL.1983,
Etchebest.Lavery.Pullman.TheorChimActa.1982,Stone.JCTC.2005}
the van der Waals repulsive and attractive potentials, 
as well as the polarizability models, i.e., all being relatively simple and computationally
inexpensive to evaluate mathematical models. 
Due to this reason, the CT effects 
%while usually of significant and non\hyp{}negligible magnitude as compared 
%to the total interaction energy, 
are not explicitly included in any molecular mechanics force field 
developed up to date.\cite{Demerdash.Yap.Head-Gordon.AnnuRevPhysChem.2014}
%such as OPLS, Amber, CHARMM, Gromos or even polarizale AMOEBA force field. 
The only way to
incorporate the CT effects in the condensed\hyp{}phase simulations
in \emph{ab initio} manner,
apart from performing full QM electronic structure simulations
or non\hyp{}force\hyp{}field\hyp{}based 
fragmentation techniques\cite{Leverentz.Maerzke.Keasler.Siepmann.Truhlar.PCCP.2012,
Dahlke.Truhlar.JCTC.2008},
is through the second generation of the effective fragment potential 
(EFP2) method,\cite{Gordon.Smith.Xu.Slipchenko.AnnuRevPhysChem.2013,
   Nguyen.Pachter.Day.JCP.2014,
   Day.Jensen.Gordon.Webb.Stevens.Krauss.Garmer.Basch.Cohen.JCP.1996,
   Sattasathuchana.Xu.Gordon.JPCA.2019,
   Blasiak.Londergan.Webb.Cho.ACR.2017,
   Kuroki.Mori.ChemLett.2016,
   Ghosh.Cho.Choi.JPCB.2014}
developed by Jensen, Gordon and co\hyp{}workers.\cite{Jensen.JCP.2001,Li.Gordon.Jensen.JCP.2006,Xu.Gordon.JCP.2013}
EFP2 is an \emph{ab initio} force field
derived on the grounds of the first\hyp{}principles %perturbation theory
at the HF level\cite{Jensen.Gordon.MolPhys.1996,Jensen.Gordon.JCP.1998,Li.Netzloff.Gordon.JCP.2006} 
and including the intermolecular dispersion effects
by the response theory.\cite{Adamovic.Gordon.MolPhys.2005,Xu.Zahariev.Gordon.JCTC.2014}
%based on the imaginary frequency\hyp{}dependent dipole\hyp{}dipole polarizabilities.
That is to say, the total intermolecular 
interaction potential, which can also be perceived as an approximation
to SAPT at HF level with added intermolecular dispersion correction, is approximated as 
%
\begin{equation}\label{e:u-efp2}
 E^{\rm EFP2} \approx E^{\rm Coul} + E^{\rm Ex-Rep} + E^{\rm Ind} + E^{\rm Disp} + E^{\rm CT} \;,
\end{equation}
%
where $E^{\rm Coul}$ is the Coulombic interaction energy of the unperturbed charge\hyp{}density
distributions of the monomers, treated by the distributed multipole approximation
with damping to account for the charge\hyp{}penetration effects,\cite{Slipchenko.Gordon.JCC.2007}
$E^{\rm Ex-Rep}$ is the exchange\hyp{}repulsion energy originating from the Pauli exclusion
principle,\cite{Jensen.Gordon.MolPhys.1996,Jensen.Gordon.JCP.1998} 
that could be roughly understood as the Heitler\hyp{}London interaction
energy\cite{Chalasinski.Gutowski.MolPhys.1985} with $E^{\rm Coul}$ subtracted, $E^{\rm Ind}$ and $E^{\rm Disp}$
are the induction and dispersion energies obtained from the distributed polarizability
approximation,\cite{Li.Netzloff.Gordon.JCP.2006,Adamovic.Gordon.MolPhys.2005,Xu.Zahariev.Gordon.JCTC.2014} 
and finally $E^{\rm CT}$ is the CT energy,\cite{Li.Gordon.Jensen.JCP.2006,Xu.Gordon.JCP.2013} 
the focus of this work.

@@

To establish the notation within the present work, 
the formalism of one\hyp{}electron potential operators is
given first.
The one\hyp{}electron Coulomb static effective potential $v^{\rm eff}({\bf r})$
produced by a certain effective one\hyp{}electron charge density distribution $\rho^{\rm eff}({\bf r})$
is 
%
\begin{equation} \label{e:v-eff}
	v^{\rm eff}({\bf r}) = \int \frac{ \rho^{\rm eff}({\bf r}') }{ \vert {\bf r}' - {\bf r} \vert} d{\bf r}' \;,
\end{equation}
%
where ${\bf r}$ is a spatial coordinate. 
For convenience, the total effective density can be split into nuclear and electronic contributions,
%
\begin{equation} \label{e:rho-eff}
 \rho^{\rm eff}({\bf r}) = \lambda \rho^{\rm eff}_{\rm nuc}({\bf r}) + \rho^{\rm eff}_{\rm el}({\bf r}) \;,
\end{equation}
%
where $\lambda$ is a certain parameter and is assumed to be either 1 or 0 in this work.
Formally, the electronic part can be expanded in terms of an effective
bond order matrix, or one\hyp{}particle density (OED), represented in a certain basis of orbitals, $\BM{\phi}({\bf r})$,
%
\begin{equation} \label{e:d-spectral}
	\rho^{\rm eff}_{\rm el}({\bf r}) = -\sum_{\alpha\beta} P_{\alpha\beta}^{\rm eff} 
	\phi_\alpha({\bf r}) \phi_\beta^{*}({\bf r})  \;.
\end{equation}
%
Based on that, the operator form of the effective potential 
can be written as
%
\begin{equation} \label{e:oep-operator}
	\hat{v}^{\rm eff} = 
        \lambda \hat{v}_{\rm nuc} +
        %\sum_{\alpha\beta} 
	%\Ket{\alpha} 
        \int d{\bf r} \Ket{{\bf r}} 
        v^{\rm eff}_{\rm el}({\bf r})
	%\left[
	%\int
	%\hat{J}^{\rm}_{\alpha\beta}({\bf r}') 
	%\rho^{\rm eff}({\bf r}')
	%d{\bf r}' 
	%\right]
	%\Bra{\beta}  
        \Bra{{\bf r}}
\end{equation}
%
with the nuclear
%and Coulomb operators 
operator
defined by
%
%\begin{subequations}
%\begin{align}
\begin{equation} \label{e:oep-operator-nuc}
        \hat{v}_{\rm nuc} \equiv \sum_x^{\rm At}  
                     \int d{\bf r} \Ket{{\bf r}} 
                     \frac{Z_x}{\vert {\bf r} - {\bf r}_x \vert}
                     \Bra{{\bf r}} \;.
	%\hat{J}^{\rm}_{\alpha\beta}({\bf r}) f({\bf r}) &\equiv
	%f({\bf r}) 
	%\int
	%\frac{ \phi_\alpha^{*}({\bf r}') \phi_\beta({\bf r}') }{ \vert {\bf r}' - {\bf r} \vert} d{\bf r}'
\end{equation}
%\end{align}
%\end{subequations}
%
%for any one\hyp{}electron function $f({\bf r})$. 
and $\BraKet{\bf r}{\alpha} \equiv \varphi_\alpha({\bf r})$.
The matrix element of the effective potential operator
is therefore given by
%
\begin{equation} \label{e:oep-matrix-element}
	\tBraKet{\alpha}{ \hat{v}^{\rm eff} }{\beta}
	= \lambda \sum_x^{\rm At} W_{\alpha\beta}^{(x)} -
        \sum_{\gamma\delta} \BraKet{\alpha\beta}{\gamma\delta} P^{\rm eff}_{\gamma\delta}  \;,
\end{equation}
%
where 
%
\begin{equation} \label{e:oep-matrix-element-nuc}
 W_{\alpha\beta}^{(x)} = 
 Z_x \int \frac{\phi^*_\alpha({\bf r}) \phi_\beta({\bf r})}{\vert {\bf r} - {\bf r}_x \vert} d{\bf r} \;,
\end{equation}
%
$Z_x$ is the atomic number of the $x$th atom,
and the symbol `At' denotes all atoms that contribute to the effective potential.
In the above equation and throughout the work, 
%
\begin{equation} \label{e:1el-operator}
\tBraKet{\alpha}{\mathscr{O}(1)}{\beta} \equiv \int \phi^*_\alpha({\bf r}) \mathscr{O}(1) \phi_\beta({\bf r}) d{\bf r} 
\end{equation}
%
for any one\hyp{}electron operator $\mathscr{O}(1)$ 
and the ERI
is defined according to
%
\begin{equation} \label{e:eri}
	\BraKet{\alpha\beta}{\gamma\delta} \equiv
	\iint 
	\frac{ \phi_\alpha^{*}({\bf r}_1) \phi_\beta({\bf r}_1) 
	       \phi_\gamma^{*}({\bf r}_2) \phi_\delta({\bf r}_2) }{ \vert {\bf r}_1 - {\bf r}_2 \vert}
	d{\bf r}_1 d{\bf r}_2  \;.
\end{equation}
%

@@

Three unique classes of ERI's can be recognized 
based on the basis function partitioning scheme
within the system composed of two molecules (shall be $A$ and $B$
throughout the course of this work). 
They are as follows:
%
\begin{enumerate}
\item the Coulomb\hyp{}like ERI's of the type 
$\BraKet{AA}{BB}\rightarrow \BraKet{\phi_{i\in A}\phi_{k\in A}}{\phi_{j\in B}\phi_{l\in B}}$,
\item the overlap\hyp{}like ERI's of the type 
$\BraKet{AA}{AB}\rightarrow \BraKet{\phi_{i\in A}\phi_{k\in A}}{\phi_{j\in A}\phi_{l\in B}}$, and
\item the exchange\hyp{}like ERI's of the type 
$\BraKet{AB}{AB}\rightarrow \BraKet{\phi_{i\in A}\phi_{j\in B}}{\phi_{k\in A}\phi_{l\in B}}$.
\end{enumerate}
%
Only the first two classes of ERI's can be incorporated into OEP's,
while exchange\hyp{}like ERI's cannot. 
The two groups, which are listed in Table~\ref{t:oep-matrix-element-types},
are usually approximated via
expanding the OEP operator in distributed multipole (DMTP) expansion series, and integrating over
one remaining electron coordinate,\cite{Li.Gordon.Jensen.JCP.2006} i.e.,
%
\begin{multline} \label{e:coulomb-energy:dmtp-oei}
 \tBraKet{j}{\hat{v}_{\rm eff}^A}{l} \cong \\
 \left< j \left| 
                   \sum_{w\in A} \left\{
                                 T^{(0)}_{we} \hat{q}_w 
                          - {\bf T}^{(1)}_{we} \cdot \hat{\BM\upmu}_w 
            + \frac{1}{3}   {\bf T}^{(2)}_{we} :     \hat{\BM\Theta}_w 
                    - \ldots  \right\}
              \right| l \right>_{{\bf r}_e} \;.
\end{multline}
%
In the above equation, $\hat{q}_{\rm eff}^{(w)}$,
$\hat{\BM\upmu}^{(w)}_{\rm eff}$ and $\hat{\BM\Theta}^{(w)}_{\rm eff}$
are the quantum operators of effective distributed monopole (charge),
dipole moment and quadrupole moment, respectively, centered on the $w$th site at ${\bf r}_w$, 
whereas ${{\bf T}^{(d)}_{we}}$ are the so called interaction tensors of rank $d$ between ${\bf r}_w$ and ${\bf r}_e$, with the latter being the electronic coordinate.
Explicit forms of interaction tensors can be found elsewhere.\cite{Blasiak.Cho.JCP.2014}
Note that $\Ket{j}$ and $\Ket{l}$ can belong to either molecule.
In this way, ERI's are no longer needed and the computational cost reduces appreciably.
In this work, this method is however considered already too expensive
for application in the CT/HF0 energy because evaluation of
Eq.~\eqref{e:coulomb-energy:dmtp-oei} requires calculation
of electrostatic potential and electrostatic potential gradient(s) OEI's. These kind of
integrals are typically the most expensive as compared to other standard OEI's
such as overlap or kinetic energy integrals.
Therefore, Coulomb and overlap\hyp{}like OEP matrix elements
%The first type is a Coulomb\hyp{}like contribution, for which 
will be treated via more approximate and less expensive approaches
%the multipole expansion or density fitting
which are discussed next.
% can be used.
%The second type is an overlap\hyp{}like contribution, which can be approximated via the density fitting. 
%Note that the exchange\hyp{}like integrals cannot be treated with the OEP technique.

%
{
\renewcommand{\arraystretch}{1.4}
\begin{table}[b]
\caption[Types of matrix elements with OEP operators]
{{\bf Types of matrix elements with OEP operators}
}
\label{t:oep-matrix-element-types}
\begin{ruledtabular}
\begin{tabular}{lcccc}
Matrix element      &&            `Overlap-like'                &&            `Coulomb-like'               \\ 
                    && $\tBraKet{i}{\hat{v}^{A}_{\rm eff}}{j} $ && $\tBraKet{j}{\hat{v}^{A}_{\rm eff}}{l}$ \\ 
	\cline{1-5}
Partitioning scheme &&            $i\in A, j\in B$              &&               $j,l\in B$                \\
ERI class           &&            $\BraKet{AA}{AB}$             &&               $\BraKet{AA}{BB}$         \\
DF\footnotemark[1]/RI\footnotemark[2] Form    
&& $\sum_{\xi\in A}^{\rm } v^A_{i\xi} S^{AB}_{\xi j} $  
&& $\sum_{\xi\zeta\in A}^{\rm } S^{BA}_{j\xi} v^A_{\xi\zeta} S^{AB}_{\zeta l} $ \\
DMTP\footnotemark[3] Form                     
&& --  &  &  $\rho_{jl}^B \odot \rho_{\rm eff}^A$ \\
\end{tabular}
\end{ruledtabular}
%
\footnotetext[1]{Density Fitting}
\footnotetext[2]{Resolution of Identity}
\footnotetext[3]{Distributed Multipole Expansion}
%
\end{table}
}
%

@@

This scheme, which will be referred to as the EDF scheme, 
is applicable in case of matrix elements of 
$
 \tBraKet{i^A}{\hat{v}_{\rm eff}^A}{j^B}
$ type. 
 These matrix elements require ERI's of type
$\BraKet{AA}{AB}$ only.
To get the \emph{ab initio} representation of such an overlap\hyp{}like matrix element,
one can use a procedure similar to
the typical density fitting (DF) or resolution of identity (RI), both of which are nowadays widely used 
to compute electron\hyp{}repulsion integrals (ERI's) more efficiently,
and reduce computational cost of post\hyp{}Hartree\hyp{}Fock methods.\cite{Hesselmann.Jansen.Schutz.JCP.2005} Density fitting was also
applied to design \emph{ab initio} force fields.\cite{Piquemal.Cisneros.Reinhardt.Gresh.Darden.JCP.2006,Cisneros.Andres.Piquemal.Darden.JCP.2005}

@@

the following objective function\cite{Cisneros.Andres.Piquemal.Darden.JCP.2005,Piquemal.Cisneros.Reinhardt.Gresh.Darden.JCP.2006}

@@

by performing integrations of Eq.~\eqref{e:triple-integral} is very costly 

@@

$H_{i\varepsilon}$ are given by Eq.~\eqref{e:v-compl}
--- thus, the matrix elements of the OEP operator can be found

@@

$V^A_{i\xi}$ can be calculated just once and stored as effective fragment parameters.

Since, in contrast to the EDF-1 scheme, now two density fitting steps are required to
obtain OEP matrix, it will be referred to as the EDF-2 scheme.
In the limitting case, when the auxiliary basis is the same as the intermediate basis,
EDF-2 is equivalent to EDF-1.
As the theory required to define OEP's and their computational
evaluation in the fragment\hyp{}based formulation has been finally given in this Section,
the application of the OEP technique of ERI elimination is
further discussed for the evaluation of the CT/HF0 energy in the following sections.

@@

 \Bra{i} \left[ -\hat{v}^B_{nj} \Ket{j} \right]  \;,
